A radiocarbon data based evaluation of Cultural Transmission in Bronze Age burial rites of Central, Northern and North-western Europe

### 2018-12-17 article/03_Results.Rmd
improvements of "Northeastern France"

General observations: **TODO**

@@

`r a("Poland")`:
The region *Poland* completely covers the centre and west of today's Poland. However, the ^14^C data from this region come almost exclusively from the voivodships of Kajuwien-Pommern, Lodsch and Lower Silesia. The amount of data is small overall, with significantly more data from the Early Bronze Age up to about 1700calBC. Almost no information is available from the Middle and Late Bronze Ages. This applies to both the *burial type* and the *burial construction* variables, the former being slightly better documented. The relative development of *burial type* approximates the *real* conditions -- i.e. according to the classical, archaeological narrative --: body burials were actually significantly more frequent in western Poland until the Late Bronze Age, while burials did not occur until the Urnfield Period. Also before, cremation graves had been laid out again and again, but in principle the tendency recorded by Radon-B is correct. In contrast, *burial construction* is misrepresented insofar as the burial mounds also played an important role in the Middle Bronze Age, which the dataset does not appreciate.

`r a("Southern Germany")`:
The artificial district region *Southern Germany* covers Bavaria completely, but also includes most of Baden-Württemberg and Thuringia as well as extensive areas in Moravia and Western Austria. No ^14^C data are available from the latter regions. The majority of the available data comes from the Upper Palatinate, Lower Bavaria and Swabia as well as (in Württemberg) the administrative districts of Tübingen. For the Early Bronze Age a remarkable amount of data from Southern Germany are preserved in Radon-B, but this amount decreases in the course of the Middle Bronze Age and leaves the Late Bronze Age almost without dated graves. The *burial type* variable is quite well understood -- although disproportionately much more of the early body graves were recorded than urnfield burials. This wealth of information contrasts sharply with the situation regarding the *burial construction* variable, for which almost no information from this area is available throughout the Bronze Age. If one considers the relative development calculated from Radon-B and compares it with the archaeological literature, then a good agreement can be found for *burial type*. Body burial dominated for a long time, but was then replaced by cremation from the Middle Bronze Age onwards and massively in the Late Bronze Age. The calculated proportions for *burial construction* could easily be dismissed as completely worthless due to the negligibly small amount of data, but in fact the proxy indicates the right tendency for early and middle Bronze Age: The dominance of flat graves was replaced by hill burials. The data set does not allow any statements to be made about the Late Bronze Age.

`r a("Northeastern France")`:
*Northeastern France* is a district region comprising the French administrative regions Grand-Est and Bourgogne-Franche-Comté, the east of Baden-Württemberg, Rhineland-Palatinate and Saarland, as well as Luxembourg and Wallonia in Belgium. Accumulations of ^14^C data are only available from Württemberg and towards the Paris Basin. The amount of data is low overall, but very low in the early and early Middle Bronze Age. For *burial type*, after all, limited statements are possible after this time window, *burial construction* on the other hand must remain obscure. The prediction of the relative development of these variables is correspondingly error-prone. After all, the parallel existence of cremation and body burial in early and middle Bronze Age is correctly reproduced. However, this already existed in the Early Bronze Age and certainly developed more nuanced than the proxy might suggest. Contrary to the derivation from the ^14^C ratios, flat graves were quite common in *Northeastern France*.

>>>>

General observations: **TODO**
Literature! **TODO**

@@

`r a("Poland")`:
The region *Poland* completely covers the centre and western part of today's Poland. However, the ^14^C data from this region are almost exclusively from the voivodships of Kajuwien-Pommern, Lodsch and Lower Silesia. The amount of data is small overall, with significantly more data from the Early Bronze Age up to about 1700calBC. Almost no information is available from the Middle and Late Bronze Ages. This applies to both the *burial type* and the *burial construction* variables, the former being slightly better documented. The relative development of *burial type* approximates the *real* conditions -- i.e. according to the classical, archaeological narrative: inhumations were actually significantly more frequent in western Poland until the Late Bronze Age, while cremations did not occur until the Urnfield Period. Also before, cremation graves had been created sometimes, but in general the tendency recorded by Radon-B is correct. In contrast, *burial construction* is misrepresented: burial mounds also played an important role in the Middle Bronze Age.

`r a("Southern Germany")`:
The artificial district region *Southern Germany* covers Bavaria completely, but also includes most of Baden-Württemberg and Thuringia as well as extensive areas in Moravia and Western Austria. No ^14^C data are available from the latter regions. The majority of the available data comes from the Upper Palatinate, Lower Bavaria and Swabia as well as (in Württemberg) the administrative district of Tübingen. For the Early Bronze Age a remarkable amount of data from Southern Germany are preserved in Radon-B, but this amount decreases in the course of the Middle Bronze Age and leaves the Late Bronze Age almost without dated graves. The *burial type* variable is quite well understood -- although disproportionately much more of the early body graves were recorded than urnfield burials. This wealth of information contrasts sharply with the situation regarding the *burial construction* variable, for which almost no information from this area is available throughout the Bronze Age. If one considers the relative development calculated from Radon-B and compares it with the archaeological literature, then a good agreement can be found for *burial type*. Inhumation dominated for a long time, but was then replaced by cremation from the Middle Bronze Age onwards and massively in the Late Bronze Age. The calculated proportions for *burial construction* could easily be dismissed as completely worthless due to the negligibly small amount of data, but in fact the proxy indicates the right tendency for early and middle Bronze Age: The dominance of flat graves was replaced by mounds. The data set does not allow any statements about the Late Bronze Age.

`r a("Northeastern France")`:
*Northeastern France* comprises the French administrative regions Grand-Est and Bourgogne-Franche-Comté, the east of the German Baden-Württemberg, Rhineland-Palatinate and Saarland, as well as Luxembourg and Wallonia in Belgium. Accumulations of ^14^C data are only available from Württemberg and towards the Paris Basin. The amount of data is low overall, but very low in the early and early Middle Bronze Age. For *burial type*, after all, limited statements are possible after this time window, *burial construction* on the other hand must remain obscure. The prediction of the relative development of these variables is correspondingly error-prone. Still the parallel existence of cremation and body burial in early and middle Bronze Age is correctly reflected. However, this situation already existed in the Early Bronze Age and certainly developed more nuanced than the proxy might suggest. Contrary to the derivation from the ^14^C ratios, flat graves were quite common in Northeastern France.
